---
title: "March Madness Model"
output: html_notebook
---
```{r}
library(Stat2Data)

mm.df <- read.csv("MarchMadness.csv")
```

```{r}
mm.glm <- glm(Win ~ PTS + FTR + OR + DR + FGA + FTA + TO + TOR, family = "binomial", data = mm.df)
summary(mm.glm)
```
We can determine that the amount of field goal attempts is not a good predictor for whether the 
team wins or not because its p-value is 0.954199 > 0.05.

```{r}
mm.glm2 <- glm(Win ~ PTS + FTR + OR + DR + FTA + TO + TOR, family = "binomial", data = mm.df)
summary(mm.glm2)
```
We can determine that the amount of free throw rate is not a good predictor for whether the 
team wins or not because its p-value is 0.0.215868 > 0.05.

```{r}
mm.glm3 <- glm(Win ~ PTS + OR + DR + FTA + TO + TOR, family = "binomial", data = mm.df)
summary(mm.glm3)
```
The offensive and defensive rebound, free throw attempts, total turnovers, and turnover rate
are good predictors of whether a team will win because their p-values are less than 0.05.

Prediction Equation:
log(Wins) = -17.58457 + (0.09994)Points + (0.18493)Offensive Rebound + (0.24748)Defensive Rebound + (0.06575)Free Throw Attempts - (0.96550)Total Turnovers + (0.81709)Turnover Rate


# Randomization

## PTS
```{r}
coefPTS <- numeric(1000)
realPTS <- coef(mm.glm)["PTS"]
for(i in 1:1000) {
  shufflePTS <- sample(mm.df$PTS)
  shufPTS.glm <- glm(Win ~ shufflePTS + FTR + OR + DR + FGA + FTA + TO + TOR, family = "binomial", data = mm.df)
  coefPTS[i] <- coef(shufPTS.glm)["shufflePTS"]
}

randPval <- mean(abs(coefPTS) >= abs(realPTS))
randPval

```
p value for randomization test for p value is so small that is approximated as 0, so PTS is a significant predictor

## FTR
```{r}
coefFTR <- numeric(1000)
realFTR <- coef(mm.glm)["FTR"]
for(i in 1:1000) {
  shuffleFTR <- sample(mm.df$FTR)
  shufFTR.glm <- glm(Win ~ PTS + shuffleFTR + OR + DR + FGA + FTA + TO + TOR, family = "binomial", data = mm.df)
  coefFTR[i] <- coef(shufFTR.glm)["shuffleFTR"]
}

randPval <- mean(abs(coefFTR) >= abs(realFTR))
randPval

```
FTR p value = 0.178, greater than 0.05, likely not significant

## OR
```{r}
coefOR <- numeric(1000)
realOR <- coef(mm.glm)["OR"]
for(i in 1:1000) {
  shuffleOR <- sample(mm.df$OR)
  shufOR.glm <- glm(Win ~ PTS + FTR + shuffleOR + DR + FGA + FTA + TO + TOR, family = "binomial", data = mm.df)
  coefOR[i] <- coef(shufOR.glm)["shuffleOR"]
}

randPval <- mean(abs(coefOR) >= abs(realOR))
randPval

```
p value <0.05, OR is significant

## DR
```{r}
coefDR <- numeric(1000)
realDR <- coef(mm.glm)["DR"]
for(i in 1:1000) {
  shuffleDR <- sample(mm.df$DR)
  shufDR.glm <- glm(Win ~ PTS + FTR + OR + shuffleDR + FGA + FTA + TO + TOR, family = "binomial", data = mm.df)
  coefDR[i] <- coef(shufDR.glm)["shuffleDR"]
}

randPval <- mean(abs(coefDR) >= abs(realDR))
randPval
```
p value <0.05, DR is significant

## FGA
```{r}
coefFGA <- numeric(1000)
realFGA <- coef(mm.glm)["FGA"]
for(i in 1:1000) {
  shuffleFGA <- sample(mm.df$FGA)
  shufFGA.glm <- glm(Win ~ PTS + FTR + OR + DR + shuffleFGA + FTA + TO + TOR, family = "binomial", data = mm.df)
  coefFGA[i] <- coef(shufFGA.glm)["shuffleFGA"]
}

randPval <- mean(abs(coefFGA) >= abs(realFGA))
randPval
```
pval=0.628 > 0.05, not significant

## FTA
```{r}
coefFTA <- numeric(1000)
realFTA <- coef(mm.glm)["FTA"]
for(i in 1:1000) {
  shuffleFTA <- sample(mm.df$FTA)
  shufFTA.glm <- glm(Win ~ PTS + FTR + OR + shuffleFTA + FGA + FTA + TO + TOR, family = "binomial", data = mm.df)
  coefFTA[i] <- coef(shufFTA.glm)["shuffleFTA"]
}

randPval <- mean(abs(coefFTA) >= abs(realFTA))
randPval
```
pval = 0.005 < 0.05, significant

## TO
```{r}
coefTO <- numeric(1000)
realTO <- coef(mm.glm)["TO"]
for(i in 1:1000) {
  shuffleTO <- sample(mm.df$TO)
  shufTO.glm <- glm(Win ~ PTS + FTR + OR + shuffleTO + FGA + FTA + TO + TOR, family = "binomial", data = mm.df)
  coefTO[i] <- coef(shufTO.glm)["shuffleTO"]
}

randPval <- mean(abs(coefTO) >= abs(realTO))
randPval
```

pval<0.05, significant

## TOR
```{r}
coefTOR <- numeric(1000)
realTOR <- coef(mm.glm)["TOR"]
for(i in 1:1000) {
  shuffleTOR <- sample(mm.df$TOR)
  shufTOR.glm <- glm(Win ~ PTS + FTR + OR + shuffleTOR + FGA + FTA + TO + TOR, family = "binomial", data = mm.df)
  coefTOR[i] <- coef(shufTOR.glm)["shuffleTOR"]
}

randPval <- mean(abs(coefTOR) >= abs(realTOR))
randPval
```
pval<0.05, significant

# BootStrap
```{r}
bootPTS <- numeric(1000)
bootFTR <- numeric(1000)
bootOR <- numeric(1000)
bootDR <- numeric(1000)
bootFGA <- numeric(1000)
bootFTA <- numeric(1000)
bootTO <- numeric(1000)
bootTOR <- numeric(1000)

for(i in 1:1000) {
  bootSample <- mm.df[sample(1:nrow(mm.df), replace=TRUE), ]
  bootModel <- glm(Win ~ PTS + FTR + OR + DR + FGA + FTA + TO + TOR, 
                      family = binomial, data = bootSample)
  bootPTS[i] <- exp(coef(bootModel)["PTS"])
  bootFTR[i] <- exp(coef(bootModel)["FTR"])
  bootOR[i]  <- exp(coef(bootModel)["OR"])
  bootDR[i]  <- exp(coef(bootModel)["DR"])
  bootFGA[i] <- exp(coef(bootModel)["FGA"])
  bootFTA[i] <- exp(coef(bootModel)["FTA"])
  bootTO[i]  <- exp(coef(bootModel)["TO"])
  bootTOR[i] <- exp(coef(bootModel)["TOR"])

}

quantile(bootPTS, c(0.025, 0.975))
quantile(bootFTR, c(0.025, 0.975))
quantile(bootOR, c(0.025, 0.975))
quantile(bootDR, c(0.025, 0.975))
quantile(bootFGA, c(0.025, 0.975))
quantile(bootFTA, c(0.025, 0.975))
quantile(bootTO, c(0.025, 0.975))
quantile(bootTOR, c(0.025, 0.975))
 
```

We are 95% confident that the odd ratio of PTS is within 1.053 to 1.182. This means each additional point scored increases the odds of winning by approximately 5.3% to 18.2%

We are 95% confident that the odd ratio of FTR is within 0.94 to 1.01. This interval includes 1, so its likely not statstically significant

We are 95% confident that the odd ratio of OR is within 1.02 to 1.37. This means each additional OR increases the odds of winning by approximately 2% to 37%

We are 95% confident that the odd ratio of DR is within 1.09 to 1.49. This means each additional DR increases the odds of winning by approximately 9% to 49%

We are 95% confident that the odd ratio of FGA is within 0.67 to 1.49. This interval includes 1, so its likely not statstically significant

We are 95% confident that the odd ratio of FTA is within 0.88 to 1.30. This interval includes 1, so its likely not statstically significant

We are 95% confident that the odd ratio of TO is within 0.04 to 3.96. This interval includes 1, so its likely not statstically significant

We are 95% confident that the odd ratio of TOR is within 0.25 to 17.9 This interval includes 1, so its likely not statstically significant

# Bracket Simulation
```{r}
br.df <- read.csv("BracketMarchMadness.csv")
logOddsOfTeams <- list(teamName = br.df$Team, logOdd = numeric(nrow(br.df)))

# for loop for setting log(Odds) for each team
for(i in seq_len(nrow(br.df))) {
    logOddsOfTeams$logOdd[i] <- mm.glm3[[1]][[1]]
    for(j in 2:7) {
      logOddsOfTeams$logOdd[i] <- logOddsOfTeams$logOdd[[i]] + br.df[[j]][[i]] * mm.glm3[[1]][[j]]
    }
}
```
This is setting up all the log(Odds) for each team.

```{r}
# Color to easily see winner
red <- "\u001b[31m"
resetColor <- "\u001b[0m"

# Function to simulate games
simulateGame <- function(team1, team2) {
  logOddsT1 <- logOddsOfTeams$logOdd[logOddsOfTeams$teamName == team1]
  logOddsT2 <- logOddsOfTeams$logOdd[logOddsOfTeams$teamName == team2]
  gameProbability <- exp(logOddsT1-logOddsT2)/(1+exp(logOddsT1-logOddsT2))

  numWins <- c(0, 0)
  for(num in runif(10000, min = 0, max = 1)) {
    if(num < gameProbability) {
      numWins[1] <- numWins[1]+1
    } 
    else {
      numWins[2] <- numWins[2]+1
    }
  }
  cat(team1, numWins[1], "\t")
  cat(team2, numWins[2], "\n")

  if(numWins[1] > numWins[2]) {
    cat(red, team1, resetColor, "\n")
    return(team1)
  }
  cat(red, team2, resetColor, "\n")
  return(team2)
}

# Special Cases (8 team duke out)
winner18 <- simulateGame(br.df$Team[18], br.df$Team[19])
winner43 <- simulateGame(br.df$Team[43], br.df$Team[44])
winner52 <- simulateGame(br.df$Team[52], br.df$Team[53])
winner61 <- simulateGame(br.df$Team[61], br.df$Team[62])

# Round 1
winnersR1 <- numeric(32)
index <- 1
for(i in 1:32) {
  if((index+1) %in% c(18, 43, 52, 61)) {
    winner <- winner61
    if(index+1 == 18) {
      winner <- winner18
    }
    else if (index+1 == 43) {
      winner <- winner43
    }
    else if (index+1 == 52) {
      winner <- winner52
    }
    winnersR1[i] <- simulateGame(br.df$Team[index], winner)
    index <- index+3
  }
  else {
    winnersR1[i] <- simulateGame(br.df$Team[index], br.df$Team[index+1])
    index <- index+2
  }
}

# Round 2
winnersR2 <- numeric(16)
index <- 1
for(i in 1:16) {
  winnersR2[i] <- simulateGame(winnersR1[index], winnersR1[index+1])
  index <- index+2
}

# Round 3
winnersR3 <- numeric(8)
index <- 1
for(i in 1:8) {
  winnersR3[i] <- simulateGame(winnersR2[index], winnersR2[index+1])
  index <- index+2
}

# Round 4
winnersR4 <- numeric(4)
index <- 1
for(i in 1:4) {
  winnersR4[i] <- simulateGame(winnersR3[index], winnersR3[index+1])
  index <- index+2
}

# Semi-Finals
winnersSF <- numeric(2)
index <- 1
for(i in 1:2) {
  winnersSF[i] <- simulateGame(winnersR4[index], winnersR4[index+1])
  index <- index+2
}

# Finals
winner <- simulateGame(winnersSF[1], winnersSF[2])
print(winner)
```