
```{r}
library(Stat2Data)

mm.df <- read.csv("MarchMadness.csv")
```

```{r}
mm.glm <- glm(Win ~ PTS + FTR + OR + DR + FGA + FTA + TO + TOR, family = "binomial", data = mm.df)
summary(mm.glm)
```
We can determine that the amount of field goal attempts is not a good predictor for whether the 
team wins or not because its p-value is 0.954199 > 0.05.

```{r}
mm.glm2 <- glm(Win ~ PTS + FTR + OR + DR + FTA + TO + TOR, family = "binomial", data = mm.df)
summary(mm.glm2)
```
We can determine that the amount of free throw rate is not a good predictor for whether the 
team wins or not because its p-value is 0.0.215868 > 0.05.

```{r}
mm.glm3 <- glm(Win ~ PTS + OR + DR + FTA + TO + TOR, family = "binomial", data = mm.df)
summary(mm.glm3)
```
The offensive and defensive rebound, free throw attempts, total turnovers, and turnover rate
are good predictors of whether a team will win because their p-values are less than 0.05.

Prediction Equation:
Win = (0.09994)Points + (0.18493)Offensive Rebound + (0.24748)Defensive Rebound + (0.06575)Free Throw Attempts - (0.96550)Total Turnovers + (0.81709)Turnover Rate

```{r}
br.df <- read.csv("BracketMarchMadness.csv")
calcValues <- numeric(nrow(br.df))

for (i in seq_len(nrow(br.df))) {
    calcValues[i] <- mm.glm3[[1]][[1]]
    for(j in 2:7) {
      calcValues[i] <- calcValues[i] + br.df[[j]][[i]] * mm.glm3[[1]][[j]]
    }
    # print(br.df[[1]][[i]])
    # print(calcValues[[i]])
}
```

```{r}
temp <- calcValues[1] - calcValues[2]
fin <- exp(temp)/(1+exp(temp))
print(fin)
# 19,44,53,62
```