---
title: "March Madness Model"
output: html_notebook
---
```{r}
library(Stat2Data)

mm.df <- read.csv("MarchMadness.csv")
mm.glm3 <- glm(Win ~ PTS + OR + DR + FTA + TO + TOR, family = "binomial", data = mm.df)
summary(mm.glm3)
```


# Bracket Simulation
```{r}
br.df <- read.csv("BracketMarchMadness.csv")
logOddsOfTeams <- list(teamName = br.df$Team, logOdd = numeric(nrow(br.df)))

# for loop for setting log(Odds) for each team
for(i in seq_len(nrow(br.df))) {
    logOddsOfTeams$logOdd[i] <- mm.glm3[[1]][[1]]
    for(j in 2:7) {
      logOddsOfTeams$logOdd[i] <- logOddsOfTeams$logOdd[[i]] + br.df[[j]][[i]] * mm.glm3[[1]][[j]]
    }
}
```
This is setting up all the log(Odds) for each team.

```{r}
# Color to easily see winner
red <- "\u001b[31m"
resetColor <- "\u001b[0m"
gameProbabilities <- numeric(40)
gP.counter <- 1

# Function to simulate games
simulateGame <- function(team1, team2, doResidual) {
  logOddsT1 <- logOddsOfTeams$logOdd[logOddsOfTeams$teamName == team1]
  logOddsT2 <- logOddsOfTeams$logOdd[logOddsOfTeams$teamName == team2]
  gameProbability <- exp(logOddsT1-logOddsT2)/(1+exp(logOddsT1-logOddsT2))

  numWins <- c(0, 0)
  for(num in runif(100000, min = 0, max = 1)) {
    if(num < gameProbability) {
      numWins[1] <- numWins[1]+1
    } 
    else {
      numWins[2] <- numWins[2]+1
    }
  }
  cat("\n", team1, ": ", numWins[1], "\t")
  cat(team2, ": ", numWins[2], "\n")

  if(doResidual) {
    gameProbabilities[gP.counter] <<- gameProbability
    gP.counter <<- gP.counter+1
  }

  if(numWins[1] > numWins[2]) {
    cat(red, team1, resetColor, "\n")
    return(team1)
  }
  cat(red, team2, resetColor, "\n")
  return(team2)
}

# Special Cases (8 team duke out)
winner18 <- simulateGame(br.df$Team[18], br.df$Team[19], FALSE)
winner43 <- simulateGame(br.df$Team[43], br.df$Team[44], FALSE)
winner52 <- simulateGame(br.df$Team[52], br.df$Team[53], FALSE)
winner61 <- simulateGame(br.df$Team[61], br.df$Team[62], FALSE)

# Round 1
winnersR1 <- numeric(32)
index <- 1
for(i in 1:32) {
  if((index+1) %in% c(18, 43, 52, 61)) {
    winner <- winner61
    if(index+1 == 18) {
      winner <- winner18
    }
    else if (index+1 == 43) {
      winner <- winner43
    }
    else if (index+1 == 52) {
      winner <- winner52
    }
    winnersR1[i] <- simulateGame(br.df$Team[index], winner, TRUE)
    index <- index+3
  }
  else {
    winnersR1[i] <- simulateGame(br.df$Team[index], br.df$Team[index+1], TRUE)
    index <- index+2
  }
}

# Round 2
winnersR2 <- numeric(16)
index <- 1
for(i in 1:16) {
  winnersR2[i] <- simulateGame(winnersR1[index], winnersR1[index+1], FALSE)
  index <- index+2
}

# Round 3
winnersR3 <- numeric(8)
index <- 1
for(i in 1:8) {
  winnersR3[i] <- simulateGame(winnersR2[index], winnersR2[index+1], FALSE)
  index <- index+2
}

# Round 4
winnersR4 <- numeric(4)
index <- 1
for(i in 1:4) {
  winnersR4[i] <- simulateGame(winnersR3[index], winnersR3[index+1], FALSE)
  index <- index+2
}

# Semi-Finals
winnersSF <- numeric(2)
index <- 1
for(i in 1:2) {
  winnersSF[i] <- simulateGame(winnersR4[index], winnersR4[index+1], FALSE)
  index <- index+2
}

# Finals
winner <- simulateGame(winnersSF[1], winnersSF[2], FALSE)
```

Because I'm lazy and don't want to split the code, here's the big chunk.

For each game, we calculate the probability of team A winning against team B in a method. 
Then, 100,000 random numbers are generated between 0 and 1 (because why not), and if it 
is lower than the calculated probability of team A winning, team A wins. If not, team B 
wins. This is counted, and then the team with more wins after going through the 100,000
numbers is the winner of the match. Then, we do this for each matchup until we have a 
winner.

# Residuals
```{r}
# Residuals for Round of 64 are already calculated.
# Residuals for Round of 32
print(gP.counter)
simulateGame("St John's", "Kansas", TRUE)

simulateGame("Arizona", "Utah St.", TRUE)
simulateGame("High Point", "Arkansas", TRUE)
simulateGame("TEXAS", "Gonzaga", TRUE)
simulateGame("Miami (FL)", "Purdue", TRUE)

simulateGame("Michigan", "Saint Louis", TRUE)
simulateGame("Kentucky", "Iowa St.", TRUE)

# Calculating average residual
residual <- numeric(39)
std.residual <- numeric(39)

for(i in 1:39) {
  if(i %in% c(2, 10, 13, 15, 18, 19, 21, 26, 35, 37, 39)) {
    residual[i] <- 0-gameProbabilities[i]
  }
  else {
    residual[i] <- 1-gameProbabilities[i]
  }
  std.residual[i] <- sqrt(gameProbabilities[i]*(1-gameProbabilities[i]))
}
print(mean(residual))
print(mean(std.residual))
```
The average residual is 0.155, and the average standardized residiual is 0.454. 
The average residual being nonzero means that our conditions for inference aren't
met because the residuals aren't normally and evenly distributed.

```{r}
cat(residual[19], "\t", std.residual[19], "\n")
```
The residual for Wisconsin vs. High Point was -0.287, meaning that the model overestimated 
Wisconsin winning.

```{r}
# Arizona vs. Utah State
cat(residual[34], "\t", std.residual[34], "\n")
```
The standardized residual is 0.341, meaning that the observed value is 0.452 standard 
deviations above the predicted value.

```{r}
print(std.residual)
```
None of the particular points are strange as they are all under 1 standard deviation 
from the predicted value.